---
title: "Assignment Multiple Linear Regression"
author: "Group 1"
format: html
editor: visual
execute:
      warning: false
---

## **Group members:**

1\) Muhammad Ruhaifi bin Muhammad Sabri (22305762)

2\) Dzul Hairy bin Mohd Ramlan (22305758)

3\) Nur Shuhaiza binti Supian (22306056)

4\) Aziemah binti Nafsun (22305774)

5\) Noor Nazli bt Din (22306061)

------------------------------------------------------------------------

## **Workflow**

-   Introduction

-   Research question

-   Methodology

-   Data Exploration

-   Perform Linear Regression Model

-   Model Selection

-   Model Diagnostic and Assumption Checking

-   Final Model Checking

-   Interpretation

------------------------------------------------------------------------

## **Introduction**

This was data from **US Health Insurance Dataset** comprise medical information and costs billed by health insurance companies. it contains six variables with 1338 observations:

1.  age: age of patient (continuous)
2.  sex: gender of the subjects (categorical : male, female)
3.  bmi: BMI of subject (continuous)
4.  children: Number of children covered by health insurance (continuous)
5.  smoker: smoking status of subject (categorical: no, yes)
6.  charges: individual medical costs billed by health insurance (continuous)

------------------------------------------------------------------------

## **Research question**

How are age, BMI, number of children, sex, and smoking status associated with health insurance costs (charges) in USD?

------------------------------------------------------------------------

## **Methodology**

We will use multiple linear regression model approach to analyse predictors to explain variation of health insurance cost in sample population. 1st model will consist of all independent variables (age, sex, BMI, children, and smoking status) and another model will be constructed by adding interaction term between age and smoking status as effect of age towards health insurance cost can be modified by individual’s smoking status. Model comparison will be executed using anova test to select the final model. Model assessment will be conducted by running the diagnostic plot and checking the model for any influential observation before final model is concluded.

------------------------------------------------------------------------

## **Data Exploration**

```{r}
#loading necessary packages
library(tidyverse)
library(haven)
library(gtsummary)
library(corrplot)
library(knitr)
library(tidyr)
library(ggplot2)
library(broom)
```

### **Load and inspect the data**

```{r}
# Load the dataset
library(readr)
insurance <- read_csv("insurance.csv")

# Inspect the data
str(insurance)
summary(insurance)
head(insurance)

# Check for missing values
sum(is.na(insurance))
```

------------------------------------------------------------------------

### **Data preparation**

Convert categorical variables (`sex` and `smoker`) to factors:

```{r}
insurance$sex <- as.factor(insurance$sex)
insurance$smoker <- as.factor(insurance$smoker)
```

Check if levels are correct:

```{r}
levels(insurance$sex)
levels(insurance$smoker)
```

### Exploratory data analysis (EDA)

**Visualise relationships:**

```{r}
pairs(insurance[, c("age", "bmi", "children", "charges")])
```

**Correlation between continuous variables:**

```{r}
cor_matrix <- cor(insurance[, c("age", "bmi", "children", "charges")])

#visualize the correlation matrix
corrplot(cor_matrix, method = "circle")
```

no obvious correlation issue between continuous variables

**Boxplots for categorical variables:**

1.  smoking status

```{r}
ggplot(insurance, aes(x = smoker, y = charges, fill = smoker)) +
  geom_boxplot() +
  labs(title = "Health Insurance Cost by Smoking Status", y = "Charges (USD)")
```

2.  gender

    ```{r}
    ggplot(insurance, aes(x = sex, y = charges, fill = sex)) +
      geom_boxplot() +
      labs(title = "Health Insurance Cost by Gender", y = "Charges (USD)")
    ```

### **Data Summary**

**Table Summary**

```{r}
insurance %>% 
  tbl_summary( label = list(age  ~ "Age", sex  ~ "Gender", bmi ~ "BMI" , children ~ "Number of Children", smoker ~ "Smoking Status",  charges ~ "Health Insurance Cost (USD)") , 
 statistic = list(all_continuous() ~ "{mean} ({sd})")) %>% bold_labels() %>% 
italicize_levels()
```

------------------------------------------------------------------------

### **Data Visualization**

For **continuous data**: We employ histograms to gain insight into the distribution of each feature. This allows us to understand the central tendency, spread, and shape of the dataset’s distribution.

1.  Age

    ```{r}
    ggplot(insurance, aes (age))+
      geom_histogram()
    ```

2.  BMI

    ```{r}
    ggplot(insurance, aes (bmi))+
      geom_histogram()
    ```

3.  Number of Children

    ```{r}
    ggplot(insurance, aes (children))+
      geom_histogram()
    ```

4.  Health Insurance Cost (USD)

    ```{r}
    ggplot(insurance, aes (charges))+
      geom_histogram()
    ```

------------------------------------------------------------------------

For **categorical data**: Bar plots are utilized to visualize the frequency of each category. This provides a clear representation of the prominence of each category within the respective feature.

1.  gender

    ```{r}
    ggplot(insurance, aes (sex))+
      geom_bar()
    ```

2.  smoking status

    ```{r}
    ggplot(insurance, aes (smoker))+
      geom_bar()
    ```

------------------------------------------------------------------------

## **Perform Linear Regression Model**

### **Univariable analysis**

1.  age

    ```{r}
    SLR_age <- lm(charges ~ age,
                  data = insurance)

    summary(SLR_age)
    ```

2.  sex

    ```{r}
    SLR_sex <- lm(charges ~ sex,
                  data = insurance)

    summary(SLR_sex)
    ```

3.  bmi

    ```{r}
    SLR_bmi <- lm(charges ~ bmi,
                  data = insurance)

    summary(SLR_bmi)
    ```

4.  number of children

    ```{r}
    SLR_children <- lm(charges ~ children,
                  data = insurance)

    summary(SLR_children)
    ```

5.  smoking status

    ```{r}
    SLR_smoking <- lm(charges ~ smoker,
                  data = insurance)

    summary(SLR_smoking)
    ```

**Table Summary for Simple Linear Regression**

```{r}
#create a summary table for each model with confidence intervals
tbl_age <- tbl_regression(SLR_age, conf.level = 0.95)

tbl_sex <- tbl_regression(SLR_sex, conf.level = 0.95)

tbl_bmi <- tbl_regression(SLR_bmi, conf.level = 0.95)

tbl_children <- tbl_regression(SLR_children, conf.level = 0.95)

tbl_smoking <- tbl_regression(SLR_smoking, conf.level = 0.95)

#Combine all tables into one
SLR_table <- tbl_stack(
  list(tbl_age, tbl_sex, tbl_children, tbl_smoking, tbl_bmi)
) %>%
  as_gt() %>%
  gt::tab_header(
    title = "Simple Linear Regression Model"
  )

#Print the summary table
SLR_table
```

### Multiple Linear Regression without Interaction (Model A)

Model A includes all covariates—age, sex, number of children, BMI, and smoking status—as predictors to examine their association with the outcome variable, health insurance charges. This approach aims to evaluate the contribution of each factor while controlling for the others in explaining variations in insurance costs.

```{r}
MLR_modelA <- lm(charges ~ age + sex + children + bmi + smoker,
                 data = insurance)

summary(MLR_modelA)
```

**Table Summary for Model A**

```{r}
tbl_regression(MLR_modelA) %>%  add_glance_table(include = c(adj.r.squared)) %>% 
  bold_labels() %>% italicize_levels() %>% add_n() %>% 
  as_gt() %>%
  gt::tab_header(title = "Multiple Linear Regression Model A",
                 subtitle = "Without Interaction")
```

### Multiple Linear Regression with Interaction (Model B)

Model B is a multiple linear regression (MLR) model that includes an interaction term between age and smoking status. This decision is based on the premise that both variables—age and smoking status—have the potential to modify their combined effect on the outcome variable, health insurance charges

```{r}
MLR_modelB <- lm(charges ~ age + sex + children + bmi + smoker + age:smoker,
                 data = insurance)

summary(MLR_modelB)
```

**Table Summary for Model B**

```{r}
tbl_regression(MLR_modelB) %>%  add_glance_table(include = c(adj.r.squared)) %>% 
  bold_labels() %>% italicize_levels() %>% add_n() %>% 
  as_gt() %>%
  gt::tab_header(title = "Multiple Linear Regression Model B",
                 subtitle = "With Interaction")
```

------------------------------------------------------------------------

## **Model Selection**

-   Compare model fit:

1.  Adjusted R^2^

    ```{r}
    summary(MLR_modelA)$adj.r.squared
    summary(MLR_modelB)$adj.r.squared
    ```

2.  Akaike Information Criterion (AIC)

    ```{r}
    AIC(MLR_modelA)
    AIC(MLR_modelB)
    ```

-   Perform an Analysis of Variance (ANOVA) to formally test whether adding the interaction term significantly improves the model.

```{r}
anova(MLR_modelA,MLR_modelB)
```

Since p-value is 0.1317, the interaction term does not significantly improve the model, thus simpler model (MLR_modelA) is chosen for ease of interpretation.

------------------------------------------------------------------------

## **Model Diagnostic and Assumption Checking**

We took Model A (MLR model without interaction) as our preliminary model

```{r}
prelim.final.m <- lm (charges ~ age + sex + children + bmi + smoker, 
                      data = insurance)
```

**a. Diagnostics plots for residual analysis**

**Plot 1 (Linearity)**

```{r}
plot(prelim.final.m,which=1)
```

Comment: random scatter of points, thus linearity assumption is likely meet.

**Plot 2 (normality of residuals)**

```{r}
plot(prelim.final.m,which=2)
```

Comment: The Q-Q plot indicates that the majority of the data points closely follow the diagonal line, suggesting that the residuals are approximately normally distributed. However, a small number of observations (specifically observations 243, 578, and 1301) exhibit large residuals, which may suggest potential outliers or influential data points.

**Plot 3 (Homoscedasticity)**

```{r}
plot(prelim.final.m,which=3)
```

Comment: residuals appear to be equally spread (random scatter) across all levels of the fitted values. thus homoscedasticity is likely valid.

Plot 4 (Influential)

```{r}
plot(prelim.final.m,which=5)
```

Comment: Observations 544, 578, and 1301 are identified as potential outliers in the Residual vs Leverage plot. However, these points do not exceed the Cook's distance threshold, indicating that they are not influential cases that would significantly impact the regression model's result.

b\. Plot residuals against numerical independent variables (eg : age, number of children, and BMI) in the model to check for individual linearity.

1.  residual vs age

```{r}
augment(prelim.final.m) %>%
  ggplot(aes(x = age, y = .resid))+
  geom_point()+
  geom_smooth()
```

2.  residual vs number of children

    ```{r}
    augment(prelim.final.m) %>%
      ggplot(aes(x = children, y = .resid))+
      geom_point()+
      geom_smooth()
    ```

3.  residual vs BMI

    ```{r}
    augment(prelim.final.m) %>%
      ggplot(aes(x = bmi, y = .resid))+
      geom_point()+
      geom_smooth()
    ```
