---
title: "Assignment Multiple Linear Regression"
author: "Group 1"
format: html
editor: visual
---

## **Group members:**

1\) Muhammad Ruhaifi bin Muhammad Sabri (22305762)

2\) Dzul Hairy bin Mohd Ramlan (22305758)

3\) Nur Shuhaiza binti Supian (22306056)

4\) Aziemah binti Nafsun (22305774)

5\) Noor Nazli bt Din (22306061)

------------------------------------------------------------------------

## **Workflow**

-   Introduction

-   Research question

-   Methodology

-   Data Exploration

-   Perform Linear Regression Model

-   Model Assessment

-   Final Model Checking

-   Interpretation

------------------------------------------------------------------------

## **Introduction**

This was data from **US Health Insurance Dataset** comprise

------------------------------------------------------------------------

## Research question

How are age, BMI, number of children, sex, and smoking status associated with health insurance costs (charges) in USD?

---

## **Methodology**

------------------------------------------------------------------------

## **Data Exploration**

```{r}
#loading necessary packages
library(tidyverse)
library(haven)
library(gtsummary)
library(corrplot)
library(knitr)
library(tidyr)
library(ggplot2)
library(broom)
```

### **Load and inspect the data**

```{r}
# Load the dataset
library(readr)
insurance <- read_csv("insurance.csv")

# Inspect the data
str(insurance)
summary(insurance)
head(insurance)

# Check for missing values
sum(is.na(insurance))
```

------------------------------------------------------------------------

### Data preparation

Convert categorical variables (`sex` and `smoker`) to factors:

```{r}
insurance$sex <- as.factor(insurance$sex)
insurance$smoker <- as.factor(insurance$smoker)
```

Check if levels are correct:

```{r}
levels(insurance$sex)
levels(insurance$smoker)
```

### Exploratory data analysis (EDA)

Visualise relationships:

```{r}
pairs(insurance[, c("age", "bmi", "children", "charges")])
```

Correlation between continuous variables:

```{r}
cor(insurance[, c("age", "bmi", "children", "charges")])
```

Boxplots for categorical variables:

```{r}
ggplot(insurance, aes(x = smoker, y = charges, fill = smoker)) +
  geom_boxplot() +
  labs(title = "Health Insurance Cost by Smoking Status", y = "Charges (USD)")
```

## **Data Summary**

```{r}
summary(insurance)
```

**Table Summary**

```{r}
insurance %>% 
  tbl_summary( label = list(age  ~ "Age", sex  ~ "Gender", bmi ~ "BMI" , children ~ "Number of Children", smoker ~ "Smoking Status",  charges ~ "Health Insurance Cost (USD)") , 
 statistic = list(all_continuous() ~ "{mean} ({sd})")) %>% bold_labels() %>% 
italicize_levels()
```

------------------------------------------------------------------------

## **Explore Data**

For **continuous data**: We employ histograms to gain insight into the distribution of each feature. This allows us to understand the central tendency, spread, and shape of the dataset’s distribution.

1.  Age

    ```{r}
    ggplot(insurance, aes (age))+
      geom_histogram()
    ```

2.  BMI

    ```{r}
    ggplot(insurance, aes (bmi))+
      geom_histogram()
    ```

3.  Number of Children

    ```{r}
    ggplot(insurance, aes (children))+
      geom_histogram()
    ```

4.  Health Insurance Cost (USD)

    ```{r}
    ggplot(insurance, aes (charges))+
      geom_histogram()
    ```

------------------------------------------------------------------------

For **categorical data**: Bar plots are utilized to visualize the frequency of each category. This provides a clear representation of the prominence of each category within the respective feature.

1.  gender

    ```{r}
    ggplot(insurance, aes (sex))+
      geom_bar()
    ```

2.  smoking status

    ```{r}
    ggplot(insurance, aes (smoker))+
      geom_bar()
    ```

------------------------------------------------------------------------

## Perform Linear Regression

### Univariable analysis

1.  age

    ```{r}
    SLR_age <- lm(charges ~ age,
                  data = insurance)

    summary(SLR_age)
    ```

2.  sex

    ```{r}
    SLR_sex <- lm(charges ~ sex,
                  data = insurance)

    summary(SLR_sex)
    ```

3.  bmi

    ```{r}
    SLR_bmi <- lm(charges ~ bmi,
                  data = insurance)

    summary(SLR_bmi)
    ```

4.  number of children

    ```{r}
    SLR_children <- lm(charges ~ children,
                  data = insurance)

    summary(SLR_children)
    ```

5.  smoking status

    ```{r}
    SLR_smoking <- lm(charges ~ smoker,
                  data = insurance)

    summary(SLR_smoking)
    ```

Table Summary

```{r}
#create a summary table for each model with confidence intervals
tbl_age <- tbl_regression(SLR_age, conf.level = 0.95)

tbl_sex <- tbl_regression(SLR_sex, conf.level = 0.95)

tbl_bmi <- tbl_regression(SLR_bmi, conf.level = 0.95)

tbl_children <- tbl_regression(SLR_children, conf.level = 0.95)

tbl_smoking <- tbl_regression(SLR_smoking, conf.level = 0.95)

#Combine all tables into one
SLR_table <- tbl_stack(
  list(tbl_age, tbl_sex, tbl_children, tbl_smoking, tbl_bmi)
) %>%
  as_gt() %>%
  gt::tab_header(
    title = "Simple Logistic Regression Model"
  )

#Print the summary table
SLR_table
```

### Multiple Linear Regression without Interaction (Model A)

```{r}
MLR_modelA <- lm(charges ~ age + sex + children + bmi + smoker,
                 data = insurance)

summary(MLR_modelA)
```

Table Summary for Model A

```{r}
tbl_regression(MLR_modelA) %>%  add_glance_table(include = c(adj.r.squared)) %>% 
  bold_labels() %>% italicize_levels() %>% add_n() %>% 
  as_gt() %>%
  gt::tab_header(title = "Multiple Linear Regression Mode A",
                 subtitle = "Without Interaction")
```

### Multiple Linear Regression with Interaction (Model B)

```{r}
MLR_modelB <- lm(charges ~ age + sex + children + bmi + smoker + age:smoker,
                 data = insurance)

summary(MLR_modelB)
```

Table Summary for Model B

```{r}
tbl_regression(MLR_modelB) %>%  add_glance_table(include = c(adj.r.squared)) %>% 
  bold_labels() %>% italicize_levels() %>% add_n() %>% 
  as_gt() %>%
  gt::tab_header(title = "Multiple Linear Regression Model B",
                 subtitle = "With Interaction")
```

------------------------------------------------------------------------

## **Model Selection**

Model comparison using anova

```{r}
anova(MLR_modelA,MLR_modelB)
```

Since p-value is 0.1317, the interaction term does not significantly improve the model, thus simpler model (MLR_modelA) is chosen for ease of interpretation.

------------------------------------------------------------------------

## **Model Assessment**

We took Model A (MLR model without interaction) as our preliminary model

```{r}
prelim.final.m <- lm (charges ~ age + sex + children + bmi + smoker, 
                      data = insurance)
```

a\. Diagnostics plots

Plot 1

```{r}
plot(prelim.final.m,which=1)
```

Comment:

Plot 2

```{r}
plot(prelim.final.m,which=2)
```

Comment:

Plot 3

```{r}
plot(prelim.final.m,which=3)
```

Comment:

Plot 4

```{r}
plot(prelim.final.m,which=5)
```

Comment:

b\. Plot residuals against numerical independent variables (eg : age, number of children, and BMI) in the model to check for individual linearity.

1.  residual vs age

```{r}
augment(prelim.final.m) %>%
  ggplot(aes(x = age, y = .resid))+
  geom_point()+
  geom_smooth()
```

2.  residual vs number of children

    ```{r}
    augment(prelim.final.m) %>%
      ggplot(aes(x = children, y = .resid))+
      geom_point()+
      geom_smooth()
    ```

3.  residual vs BMI

    ```{r}
    augment(prelim.final.m) %>%
      ggplot(aes(x = bmi, y = .resid))+
      geom_point()+
      geom_smooth()
    ```
